Interactive Summarization of Relational Tables

### 2014-10-25 TableSummarization/Paper/TableSummarization.tex
Added Sampling scheme

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints. The summary is in form of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of the subset. We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table. Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
% TODO: Improve abstract.
\end{abstract}

@@

Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it. Large amounts of data hard for human to make sense of. This necessitates summarizing the data in an informative, interpretable way. The summary also needs to be fast and interactive to facilitate deeper exploration.

@@

\begin{example}\label{ex:introexample}
Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of \textit{rules}, with the \textit{count} denoting the number of tuples that satisfy the rule, and the \textit{size} denoting the number of non-$\star$ values of the rule) for each rule.

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
target & bicycles & $\star$ & $200$ & $1$ \\ \hline
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

The first rule tells the user that there are $6000$ tuples in all. 
The next rule says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles. The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores. The last rule tells the user that walmart is doing well in general over multiple products and states.

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
target & bicycles & $\star$ & $200$ & $1$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}

This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.
\end{example}

The above example raises some questions regarding how to score the rules, and what operations to allow the user to perform to dig deeper into the rules.
We describe our user interface in Section \ref{sec:interface}. Regarding scoring, we generally want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible. Thus scoring rules involves a tradeoff between specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$). We further discuss our scoring scheme in Section~\ref{sec:weighting}.

There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}. This work mostly focuses on giving he most 'surprising' information to the user i.e. information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution. Our algorithm, on the other hand, focuses on coverage, describing as much of the table as possible.

%??Something about how users can't imagine max-entropy distribution??.

OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'. We focus on only obtaining counts for rules which cover a significant number of tuples, to get a fast response time. We compute more detailed rules on the fly, when the user explores the table further. In addition, we often display rules containing multiple non-$\star$ values in our summary (such as the (target, bicycles) rule in our example) is their score is high enough, whereas a user would need to multiple clicks in an OLAP system to get the same information. Finally, our algorithm has the flexibility of allowing the user to specify a rule-weighting function, in order to preferentially choose rules of a particular size, or containing information about a particular column that the user is interested in.

% TODO : Need motivation for coverage over surprise. Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
\stitle{Overview of paper:} 
\squishlist

@@

\section{Preliminaries} 
\label{sec:preliminaries}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table. Our objective (formally defined later) is to summarise the table with a list of {\em rules}. A {\em rule} is a tuple with one field per column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each field of the rule can either be one of the values in the corresponding column of the table, or $\star$, which is a wildcard character representing all values in the column. For a table column with numerical values, we can allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all fields. A rule-list is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user. A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

@@

When a user clicks on a rule to know more about the part of the table covered by that rule, we display a list of sub-rules of the rule. This list is called a {\em rule-list}.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule. {\em Count}($r$) is defined as the total number of tuples $t \in T$ that are covered by $r$. Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list. Let {\em Size}($r$) refer to the number of non-starred values in rule $r$. We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} \text{MCount}(r)W(r)$$.

@@

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r$, we want to find the highest scoring list of rules to expand rule $r$ into. We now formally state the problem:

Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ rules that maximizes the total score given by :
$$\sum_{r \in R}W(r)\text{MCount}(r)$$

$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily understandable summary of the table. We now describe a greedy algorithm for finding the approximately highest scoring rule-list.

The process of finding the rule-list can be broken down into two steps: Finding the set of rules to put in the list, and then ordering them. We first note that once the set of rules is given, then the optimal ordering is to order them in descending order by weight, since each tuple covered by one of the rules can contribute $1$ to the MCount of exactly one of the rules, and it can contribute the most to total score by assigning it the the rule with highest weight.

Thus the hard part of the problem is finding a set of rules that maximises total score when ordered in decreasing order by weight. We solve this part of the problem next.

\subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
Define the score of a set of rules $R$ to be equal to the score for the list obtained by sorting the set elements in decreasing order by weight. The key thing to note here is that the rule weight $W(r)$ depends on the rule alone and not on other rules or on distribution of column values in the database table. The total score is given by $\sum_{r \in R} \text{Count}(r)W(r)$ and thus rule $r$ 'contributes' MCount($r$)$W(r)$ to the total score. Thus the only part of the contribution that depends on the other rules in $R$ is MCount($r$) (the dependence is because other rules in $R$ may cover some rules that $r$ covers, causing MCount($r$) to go down). We assume that rules in the rule set are ordered in decreasing value of $W(r)$, and this a tuple is always counted for the first rule that covers it. In that case, note that if $r \in R_1 \subset R_2$, then Count($r$) in $R_1$ is going to be $\geq \text{Count}(r)$ in $R_2$. Thus the marginal value that $r$ adds to the total score is higher for $R_1$ than for $R_2$, and hence the total score for set $R$ is submodular.

\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}

Intuitively, this means that the marginal value of adding an element to a set cannot increase if we first add other elements to a set. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way. Specifically, consider the greedy algorithm:

\begin{framed}
\begin{enumerate}
\item Set $R = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($R \cup \left\lbrace r \right\rbrace$) is the highest.
\item $R = R \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\end{framed}

Because Score is a submodular function of the set $R$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm for Rule-Based Scoring Functions}. The algorithm takes four parameters as input: the table $T$, the number $n$ of rules required in the final solution list, a parameter $m_w$ which stands for \textit{Max Weight}, and the weight function $W$. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Smaller values of $m_w$ will cause the algorithm to run faster, but make it more likely to miss out on high-weight rules in the final solution.

The algorithm initializes the solution set $S$ to be empty, and then iterates for $n$ steps, adding the best marginal rule at each step. To find the best marginal rule, it calls the function in the box labelled \textbf{Find best marginal rule}.

Each call of \textbf{Find best marginal rule} needs to find the marginal counts of several rules to choose the best one. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory. In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}. We count the rule over several passes, with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts for rules of size $j$. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted so far. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous pass. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

@@

\begin{framed}
\textbf{Greedy Algorithm for Rule-Based Scoring Functions}

\textbf{Input :} $n$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)

\textbf{Output :} $S$ (Solution set of rules)
\begin{enumerate}
\item $S = \phi$ // Initialize solution set to null set.
\item For $i$ from $1$ to $n$:
\begin{enumerate}[label*=\arabic*.]
\item $R_m = \text{Find best marginal rule}(S, T, m_w, W)$
\item $S = S \cup \left\lbrace R_m \right\rbrace$
\end{enumerate}
\item Return $S$
\end{enumerate}

@@

\subsection{Sampling for big tables}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if it does not fit in main memory. If we want exact counts for rules, we have not choice but to read the entire table. But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
The $minSS$ parameter is the minimum umber of sample tuples we are required to find counts on while running our greedy algorithm, and it determines how accurate our count estimates will be. We provide a way to find reasonable values of $minSS$ later.

At all points, the SampleHandler maintains a set of Table Samples in memory. Each table sample $s$ has the following attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate counts of a rule on the sample, into estimated counts on the entire table. The sum of $|T_s|$ is not allowed to exceed capacity $M$ at any point.

Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ with $f_s = r$ and $|T_s| \geq minSS$. Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get the estimated count on the entire table.

>>>>

\begin{abstract}
In this paper, we study the following problem: Given a relational table, find the best summary of the table, subject to user-specified constraints. The summary consists of a list of `rules', where a rule is a pattern specifying a subset of tuples, along with the size of that subset. We allow user to specify what kind of rules to prefer in the summary, and to further selectively explore parts of the table. Finally, we demonstrate the usefulness of our summary by performing experiments on a real dataset.
% TODO: Improve abstract.
\end{abstract}

@@

Data analysts need to sift through large amounts of data to learn about trends and obtain insights from it. Large amounts of data are hard for a human to make sense of. This necessitates summarizing the data in an informative, interpretable way. The summary also needs to be fast and interactive to facilitate deeper exploration.

@@

\begin{example}\label{ex:introexample}
% Need justification for how investor got data. Or change example.
Consider a table with columns `Department Store', `Product', `State' and `Number of Sales'. Suppose an investor wants to query for tuples where Sales were higher than some threshold, in order to decide where to invest his money. The resulting query can have a very large number of tuples, making it hard for a user to get valuable information out of it. We propose to summarize the result of the query with a set of `rules'. For example, a possible rule summarization for the query is shown in table~\ref{table:introexample}. The summary consists of a list of `rules', with the `count' denoting the number of tuples that satisfy the rule, and the `size' denoting the number of non-$\star$ values of the rule for each rule.

%TODO: Change indentation format. All indentation at start, number of indents equal to depth of rule
\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\ \hline
target & bicycles & $\star$ & $200$ & $2$ \\ \hline
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \hline
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample}}
\end{table}

% Things to work into example : Number of rules. Ordering.

The first rule tells the user that there are $6000$ tuples in all. We call it an `empty' rule, because it has only $\star$ values. The three additional rules give more information about the table. The number of additional rules ($3$ in this case) is a parameter that can be set by the user. The rules are shown in decreasing order by size.

The second rule in the table says that there are $200$ tuples with target as the first column value and bicycle as the second, which tells the user that target is selling a lot of bicycles. The rule after that tells the user that comforters are selling a lot in Massachusetts, across multiple department stores. The last rule tells the user that walmart is doing well in general over multiple products and states.

@@

\begin{table}
\centering
\begin{tabular}{| l | l | l | l | l |}
\hline Store & Product & State & Count & Size \\
\hline
$\star$ & $\star$ & $\star$ & $6000$ & $0$ \\  \cline{1-5}
target & bicycles & $\star$ & $200$ & $2$ \\ \cline{1-5}
$\star$ & comforters & Massachusetts & $600$ & $2$ \\ \cline{1-5}
walmart & $\star$ & $\star$ & $1000$ & $1$ \\ \cline{2-5}
$\triangleright$ walmart & cookies & $\star$ & $200$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & California & $150$ & $2$ \\ \cline{2-5}
$\triangleright$ walmart & $\star$ & Washington & $130$ & $2$ \\ \hline
\end{tabular}
\caption{Example of a rule based table summary \label{table:introexample2}}
\end{table}
% TODO: Should we display weight rather than size in table? After all, size is obvious. In intro example, we say that for this table, weight equals number of non-$\star$ columns.

The three new rules in this table give the user additional information about the $1000$ walmart tuples. This table tells the user that walmart sells a lot of cookies and that it has the most sales in California and Washington, in addition to the information given by the rules in the previous table.

In our summary, we want our rules to be descriptive by having fewer $\star$ values, but at the same time, we want our limited set of rules to cover as many tuples as possible. Thus choosing rules involves a tradeoff between highly specific rules like (walmart, cookies, California, $30$, $3$) and highly general rules like (walmart, $\star$, $\star$, $1000$, $1$). Our decision on what rules to choose is influenced by a rule-weighting scheme, which is further discussed in Section~\ref{sec:weighting}. Briefly, we allow the user to specify a weighting function $W$ as a black box, and display the optimal set of rules as per that weighting function.

This example also raises questions on what operations to allow the user to perform to dig deeper into the rules. We describe our user interface in Section \ref{sec:interface}. 
\end{example}

To summarize, we attempt to summarize relational data in a way that is:
\squishlist
\item \textbf{Flexible : } We allow users to specify a weighting function, which can be used among other things, to express preferences for rules of a certain size, or for rules containing values from some specific column.
\item \textbf{Interactive : } We have a user interface that allows a user to dig deeper into certain parts of the table.
\item \textbf{Online : } Does not require any heavy pre-computation. 
\squishend

There has been some work on finding anomalies in a dataset~\cite{Sarawagi:2001:UMA:767141.767148, Sarawagi00user-adaptiveexploration, Sarawagi98discovery-drivenexploration, DBLP:journals/pvldb/GebalyAGKS14}. These papers mostly focuses on giving the most 'surprising' information to the user i.e. information that would minimize the Kullback-Liebler(KL) divergence of the resulting max entropy distribution. For instance, if a certain set of values occur together in an unexpectedly small number of tuples, that set of values may be displayed to the user. In contrast, our algorithm focuses on rules with high counts and low intersection, covering as much of the table as possible.
% TODO: Say it's complementary. Also give an example where your thing is good. Mention MDL here. Or that max-entropy is hard to interpret for human, even though formally very clear. Our thing also works well when user doesn't know the set of column values. The set, at least number of column values, are needed for imagining the max-entropy distribution.

OLAP (Online-Analytic-Processing) systems attempt to summarize a dataset by computing counts for all `rules'. They have an operation called drill-down, which allows a user to get counts for rules having values in a given column. For instance in our example, a user can first drill down along the `Department Store' column, which will show the total number of tuples for each Department Store. After that, the user may drill down along the product column for `Walmart', which will show the user the number of tuples for each product in Walmart. Thus the user is required to decide which columns to drill down along in order to get interesting or frequent rules, and requires $k$ drill downs to get to rules of size $k$. In contrast, our summary may immediately display a size $k$ rule to the user if it has sufficiently high weight and count. There is no notion of ranking or ordering of rules in OLAP, since all possible rules are displayed for any drill down. In contrast, we only display the top few rules in terms of their score, and display them in decreasing order by weight. This makes the summary shorter and more readable by focusing only on the frequently occurring values in columns.

% TODO : Also motivate coverage based scoring over only giving top scoring rules i.e. MCount over Count. 
\stitle{Overview of paper:} 
\squishlist

@@

\section{Preliminaries} 
\label{sec:preliminaries}
Suppose we have a relational table, and we let $T$ denote the set of tuples in the table, and $C$ denote the set of columns of the table. Our objective (formally defined later) is to summarise the table with a list of {\em rules}.

A {\em rule} is a tuple with one field per column of the queried table. In addition, a rule has other attributes, such as count and weight (which we define later) associated with it. The value in each field of the rule can either be one of the values in the corresponding column of the table, or $\star$, which is a wildcard character representing all values in the column. For a table column with numerical values, we can allow the corresponding rule-value to be a range instead of a single value. A {\em trivial rule} is one that has a $\star$ value in all fields. A rule $r$ is said to {\em cover} a tuple  $t$ from the table if all non-$\star$ values in the columns of the rule match the corresponding values in the tuple. We abuse notation to write this as $t \in r$.

A {\em rule-list} is an ordered list of rules returned by our system as a summary of the table, or a part of the table specified by the user.

@@

When a user clicks on a rule to know more about the part of the table covered by that rule, we display a rule-list of sub-rules of the rule.
For instance, the second, third and fourth rule from Table~\ref{table:introexample} form a rule-list, which is displayed when the user expands the first (trivial) rule. Similarly, the second, third and fourth rules in Table~\ref{table:introexample2} form a rule-list, as do the fifth, sixth and seventh rules.

Let $R$ denote a rule-list. For any rule $r \in R$, we let $SEQ(r)$ denote the position of $r$ in the list $R$. 
Thus the $SEQ$ value of the cookies rule in Table~\ref{table:introexample2} is $1$, since it is the first rule in the list of rules obtained by expanding the walmart rule. {\em Count}($r$) is defined as the total number of tuples $t \in T$ that are covered by $r$. Let {\em MCount}($r$) (which stands for `Marginal Count') be the number of tuples covered by $r$ but not by any rule before $r$ in the rule-list $R$. A high value of MCount indicates that the rule not only covers a lot of tuples, but also covers parts of the table not covered by previous rules. We want to pick rules with a high value of MCount to display to the user, to increase the coverage of the rule-list. Let {\em Size}($r$) refer to the number of non-starred values in rule $r$. We let $W$ denote a function that assigns a {\em weight} to a rule based on how good the rule is, with higher weights assigned to better rules. A weighting function is {\em monotonic} if for all rules $r_1$, $r_2$ such that $r_1$ is a sub-rule of $r_2$, we have $W(r_1) \leq W(r_2)$. We further describe our weighting functions in Section~\ref{sec:weighting}. The total score for our list of rules is given by $$\text{Score}(R) = \sum_{r \in R} \text{MCount}(r)W(r)$$.

@@

\section{Algorithms}
\label{sec:algorithms}
When the user expands a rule $r^{\prime}$, we want to find the highest scoring list of rules to expand rule $r^{\prime}$ into. We now formally state the problem:

Given a table $T$, a weight function $W$, and a number $k$, find the list $R$ of $k$ sub-rules of $r^{\prime}$ that maximizes the total score given by :
$$\text{Score}(R) = \sum_{r \in R}W(r)\text{MCount}(r)$$

$W$ is required to be non-negative and monotonic. $k$ will usually be small (though the correctness of our solution does not depend on this), since the $k$ rules generated should constitute an easily visualizable summary of the table.

We will describe a greedy algorithm for finding the approximately highest scoring rule-list. First, we show that the rules in the optimal list must effectively be ordered in decreasing order by weight.

\begin{lemma}\label{lemma:rule-ordering}
Let $R$ be a rule-list. Let $R^{\prime}$ be the rule-list having the same rules as $R$, but ordered in descending order by weight. Then
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
\end{lemma}
\begin{proof}
We say a tuple $t$ is {\em assigned} to a rule $r$ in a rule-list $R$ if $r$ is the first rule in $R$ to cover $t$. For any rule-list $R$, let $W_{R}(t)$ be the weight of the rule in $R$ that $t$ is assigned to. Then Score can be rewritten as:
$$\text{Score}(R) = \sum_{t\in T}W_{R}(t)$$
Thus each tuple $t$ `contributes' $W(t)$ to the total score. Since $R^{\prime}$ has rules ordered in descending order by weight, the first rule in $R^{\prime}$ to cover $t$ must be the highest-weighted rule in $R$ to cover $t$. Thus $\forall t \in T : W_{R^{\prime}}(t) \geq W_{R}(t) $, which gives us
$$\text{Score}(R^{\prime}) \geq \text{Score}(R)$$
as required.
\end{proof}

Thus it is sufficient to restrict our attention to rule-lists that have rules sorted in decreasing order by weight. Or equivalently, given any set $S$ of rules, we can define the Score$(S)$ to be the score of the list obtained by sorting elements of $S$ in decreasing weight order. We want to find the $S$ that maximises $\text{Score}(S)$. We solve this part of the problem next.

\subsection{Greedy algorithm for finding optimal rule-set to summarize of a table}
We will now show that the Score function over sets of rules has a property called submodularity, which gives us a greedy approximation algorithm for optimizing it. 
\begin{definition}
A function $f: 2^S \rightarrow \mathbb{R}$ for any set $S$ is said to be submodular if and only if, for every $s \in S$, and $A \subset B \subset S$ with $s \notin A$:
$$f(A \cup \left\lbrace s \right\rbrace) - f(A) \geq f(B \cup \left\lbrace s \right\rbrace) - f(B)$$
\end{definition}
Intuitively, this means that the marginal value of adding an element to a set $S$ cannot increase if we add it to a superset of $S$ instead. For monotonic non-negative submodular functions, the problem of finding the set of a given size with maximum value for the function can be found approximately in a greedy way.

\begin{lemma}\label{lemma:submodular}
For a given table $T$, the Score function over sets $S$ of rules, defined by :
$$\text{Score}(S) = \sum_{r \in S} \text{MCount}(r)W(r)$$
is submodular.
\end{lemma}
\begin{proof}
For any set of rules $S$, let $W_{S}(t)$ be the weight of the highest-weighted rule in $S$ that covers $t$. Then Score can be rewritten as:
$$\text{Score}(S) = \sum_{t\in T}W_{S}(t)$$
Tuple $t$ `contributes' $W_{S}(t)$ to the Score. Now when we add a rule $r$ to $S$, the Score changes to :
$\text{Score}(S \cup \left\lbrace r \right\rbrace) = \sum_{t\in T}W_{S \cup \left\lbrace r \right\rbrace}(t)$
For any tuple $t$ not covered by $r$, the contribution of $t$ to the score does not change. If $t$ is covered by $r$, then 
$$W_{S\cup \left\lbrace r \right\rbrace}(t) = \textrm{max}(W_{S}(t), W(r))$$. Thus the contribution of $t$ increases by $\textrm{max}(0, W(r) - W_{S}(t))$. So we can write
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) = \sum_{t \in T(r)} \textrm{max}(0, W(r) - W_{S}(t))$$ where $T(r)$ is the set of tuples covered by $r$.

Now if a rule set $S^{\prime}$ is a super-set of $S$, then for all tuples $t$, we have $W_{S^{\prime}}(t) \geq W_{S}(t)$. Thus,
$\textrm{max}(0, W(r) - W_{S}(t)) \geq \textrm{max}(0, W(r) - W_{S^{\prime}}(t))$. Summing over all tuples $t \in T(r)$ gives us:
$$\text{Score}(S \cup \left\lbrace r \right\rbrace) - \text{Score}(S) \geq \text{Score}(S^{\prime} \cup \left\lbrace r \right\rbrace) - \text{Score}(S^{\prime})$$
which means that Score is submodular.
\end{proof}

Now consider the greedy algorithm:

\begin{framed}
\begin{enumerate}
\item Set $S = \phi$
\item For $i$ from $1$ to $k$
\begin{enumerate}
\item Find the rule $r$ for which Score($S \cup \left\lbrace r \right\rbrace$) is the highest.
\item $S = S \cup \left\lbrace r \right\rbrace$
\end{enumerate}
\end{enumerate}
\end{framed}

Because Score is a submodular function of the set $S$, this greedy algorithm is guaranteed to give us a score within a $1 - \frac{1}{e}$ factor of the optimum.

Our algorithm pseudo-code is given in the box labelled {\em Greedy Algorithm}. The algorithm takes four parameters as input: the table $T$, the number $k$ of rules required in the final solution list, a parameter $m_w$ (which we describe in the next paragraph), and the weight function $W$.

$m_w$ stands for \textit{Max Weight}. The parameter $m_w$ tells the algorithm to assume that all rules that get selected in the optimal solution are going to have weight $\leq m_w$. Thus, if $S_o$ denotes set of rules with maximum score, then as long as $m_w \geq \textrm{max}_{r \in S_o}W(r)$, our algorithm is guaranteed to return $S_o$. On the other hand if $m_w < W(r)$ for some $r \in S_o$, then there is a chance that the set returned by our algorithm does not contain $r$. Our algorithm runs faster for smaller values of $m_w$, and may only return a suboptimal result if $m_w \geq \textrm{max}_{r \in S_o}W(r)$. In practise, $\textrm{max}_{r \in S_o}W(r)$ is usually small. This is because as the size (and weight) of a rule increases, it's Count falls rapidly. The Count tends to decrease exponentially with rule size, while Weight increases linearly for common weight functions (such as $W(r) = \text{Size}(r)$). Thus rules with high weight and size have very low count, and are unlikely to occur in the optimal solution set $S_o$. Our experiments also show that the weights of rules in the optimal set are small.

% TODO: scheme/heuristic for finding good m_w values, since it seems arbitrary. Later, experiments testing effect of m_w on running time, and experiments to test finding heuristic, and show that values are in fact small.

The \textbf{Greedy Algorithm} initializes the solution set $S$ to be empty in Step $1$, and then iterates for $k$ steps in Step $2$, adding the best marginal rule at each step. To find the best marginal rule, it calls a function to find the best marginal rule given the existing set of rules $S$.

In order to find the best marginal rule, we need to find the marginal values of several rules and then choose the best one. A brute-force way to do this would be to enumerate all possible rules, and to find the marginal value for each of those rules in a single pass over the data. But the number of possible rules may be almost as large as the size of the table itself, making this step very expensive in terms of computation and memory.

In order to avoid counting too many rules, we use an idea from the {\em a priori} algorithm for frequent itemset mining~\cite{apriori}. We compute the best marginal rule over several passes (Step $3$), with the maximum number of passes equal to the maximum size of a rule. In the $j^{th}$ pass, we compute counts for rules of size $j$. To give an example, suppose we had three columns $c_1$, $c_2$, and $c_3$. In the first pass, we would compute the marginal values of all rules of size $1$. In the second pass, instead of finding marginal values for all size $2$ rules, we can use our knowledge of counts from the first pass to upper bound the potential counts of size $2$ rules, and be more selective about which rules to count in the second pass. For instance, suppose we know that the rule $(a, \star, \star)$ has a count of $1000$, while $(\star, b, \star)$ has a count of $100$. Then for any value $c$ in column $c_3$ we would know that the count of $(\star, b, c)$ is at most $100$ because it cannot exceed that of $(\star, b, \star)$. This implies that the maximum marginal value of any super-rule of $(\star, \star, c)$ having weight $\leq m_w$ is at most $100m_w$. If the rule $(a, \star, \star)$ has a marginal value of $800$, then the marginal value of any super-rule of $(\star, b, \star)$ cannot possibly exceed that of $(a, \star, \star)$. Since our aim is to only find the highest marginal value rule, we can skip counting for all super-rules of $(\star, b, \star)$ for future passes.

We now describe the function to find the best marginal rule. The pseudo-code for the function is in the box titled \textbf{Find best marginal rule}. The function maintains a threshold $H$, which is the highest marginal value that has been found for any rule so far. The function makes several iterations (Step $3$), counting marginal values for size $j$ rules in the $j^{th}$ iteration. We maintain three sets of rules. $C$ is the set of all rules whose marginal values have been counted so far. $C_n$ is the set of rules whose marginal values are going to be counted in the current pass. And $C_o$ is the set of rules whose marginal values were counted in the previous pass. For the first pass, we set $C_n$ to be all rules of size $1$. Then we compute marginal values for those rules, and set $C = C_o = C_n$.

@@

\begin{framed}
\textbf{Greedy Algorithm}

\textbf{Input :} $k$ (Number of rules required), $T$ (database table), $m_w$ (max weight), $W$ (weight function)

\textbf{Output :} $S$ (Solution set of rules)
\begin{enumerate}
\item $S = \phi$ // Initialize solution set to null set.
\item For $i$ from $1$ to $k$:
\begin{enumerate}[label*=\arabic*.]
\item $R_m = \text{Find best marginal rule}(S, T, m_w, W)$
\item $S = S \cup \left\lbrace R_m \right\rbrace$
\end{enumerate}
\item Return $S$
\end{enumerate}

@@

\subsection{Sampling for big tables}
Our greedy algorithm needs to make multiple passes over the entire table in order to find counts of rules. These passes can be very expensive if the table is large, especially if it does not fit in main memory. If we want exact counts for rules, we have no choice but to read the entire table.

But if we are willing to accept approximate counts rather than exact counts for rules, we can speed up our algorithm by loading a sample of the table into main memory, finding rule counts on the sample, and scaling up the count. If we had obtained the sample by sampling each tuple with probability $p$, then we must scale up the sample count of each rule by $\frac{1}{p}$ to get an estimate of it's count over the full table.

Thus, we use sampling to trade-off a small amount of accuracy for a faster response time. Our system includes a {\em SampleHandler}, which is given a certain memory capacity $M$, and a minimum sample size $minSS$ (specified by the user).
The $minSS$ parameter is the minimum number of sample tuples that are used to find counts while running our greedy algorithm, and this parameter determines how accurate our count estimates will be. We provide a way to find reasonable values of $minSS$ later.

At all points, the SampleHandler maintains a set of Samples in memory. Each sample $s$ is an object with three attributes: A `filter' rule $f_s$, a scaling factor $N_s$ and a set $T_s$ of tuples from the table. The set $T_s$ consists of a $\frac{1}{N_s}$ uniformly sampled fraction of tuples covered by $f_s$. The scaling factor $N_s$ is used to translate counts of a rule on the sample, into estimated counts on the entire table. The sum of $|T_s|$ is not allowed to exceed capacity $M$ at any point.

Whenever the user expands a rule $r$, our system calls the SampleHandler with argument $r$, which returns a sample $s$ whose filter value is given by $f_s = r$ and has $|T_s| \geq minSS$. The SampleHandler also computes $N_s$ when a sample is created. Then we run our greedy algorithm on the sample (with a modified weight function in case the user clicked on a $\star$) to obtain the list of rules to display. The counts of the rules on the sample are multiplied by $N_s$ before being displayed, to get the estimated count on the entire table.
